---
title: "Diabetes MH and PH Report -- Analyzing as Counts"
author: "Joshua Lambert, PhD"
date: "11/6/2019"
output: pdf_document
---

## Loading the data
```{r,echo=FALSE, message=FALSE,}
m3 <- read.csv("~/CON_PROJECTS/NHANES_Alc/m3_counts.csv") #file created by datapackages.R
library(survey)
library(tableone)
library(knitr)
library(nhanesA)
library(plyr)
library(rFSA)
library(sjPlot)
library(sjmisc)


#dim(m3)
m3<-m3[which(m3$RIDAGEYR>=18),] #remove those less than 18.
#dim(m3)

adjust4b<-c('RIDAGEYR','gender','race','edu','ses','BMXBMI',
            'SDMVPSU','SDMVSTRA','WTINT2YR')

nutr<-unique(c(grep(pattern = "lbx",x = tolower(colnames(m3))),grep(pattern = "lbd",x = tolower(colnames(m3)))))
candidate_vars<-c(adjust4b,colnames(m3)[nutr],"diab","mh","ph")

dat<-m3[,unique(candidate_vars)]
#dim(dat)

#factoring mh and ph as binary
#dat$mh<-factor(dat$mh)
#dat$ph<-factor(dat$ph)

#dichotomize countinuous variables 1=above median, 0=below median
medians<-lapply(X = dat[,c(10:54)],FUN = function(x){median(x,na.rm = TRUE)})
numNA<-lapply(X = dat[,c(10:54)],FUN = function(x){sum(is.na(x))})
mat<-cbind(t(t(medians)),t(t(numNA)))
kable(mat[-grep(rownames(mat),pattern = "SI"),],col.names = c("Median","NumNA"))

vita<-ifelse(dat$LBXVIA<60,yes = "Low-Average",no = "High")


dat[,c(10:54)]<-lapply(X = dat[,c(10:54)],FUN = function(x){factor(dicho(x,dich.by ="median"))})

dat$RIDAGEYR<-as.numeric(dat$RIDAGEYR)
dat$BMXBMI<-as.numeric(dat$BMXBMI)
dat$mh<-as.numeric(dat$mh)
dat$ph<-as.numeric(dat$ph)


vec<-grep(colnames(dat),pattern = "SI")
#vec<-which(apply(X = dat,MARGIN = 2,FUN = function(x){sum(is.na(x))})>8588) #columns with more than 50% data missing
dat<-dat[,-vec]
#leaves 17176 observations and 34 columns

dat$ph2<-ifelse(dat$ph>1,yes = 1,no = 0)
dat$mh2<-ifelse(dat$mh>1, yes=1, no=0)
```

\clearpage
\pagebreak

```{r, echo=FALSE}
kableone(CreateTableOne(vars=colnames(dat[,-c(7:10)]),data = dat[,-c(7:10)]),strata = c("diab"))
```

\clearpage
\pagebreak

```{r, echo=FALSE}
#ph
kableone(CreateTableOne(vars=colnames(dat[,-c(7:10,32,34,35,36)]),data = dat[,-c(7:10)],strata = c("ph2","diab")))
```

\clearpage
\pagebreak

```{r, echo=FALSE}
#mh
kableone(CreateTableOne(vars=colnames(dat[,-c(7:10,32,33,35,36)]),data = dat[,-c(7:10)],strata = c("mh2","diab")))
```



\clearpage
\pagebreak

# Mental Health
Analysis based on as low as 7110 subjects after removing all missingness.
```{r, echo=FALSE, include=FALSE}
##########################
#### Mental Health #######
##########################
nhanesdesign<-svydesign(id=~SDMVPSU,
                        strata=~SDMVSTRA,
                        weights=~WTINT2YR,
                        nest=TRUE,
                        data=dat)

null_model<-svyglm(formula = "mh~RIDAGEYR+gender+race+edu+ses+BMXBMI+ph+diab",design=nhanesdesign,family=quasipoisson,data=dat)
#race and BMI not significant will remove from null model and won't use in fixvar
null_model<-svyglm(formula = "mh~RIDAGEYR+gender+edu+ses+ph+diab",design=nhanesdesign,family=quasipoisson,data=dat)

dim(dat[complete.cases(dat[,-c(7:10)]),])

###########
#2-Way Int#
###########
fsafit_mh_2<-FSA(formula = "mh~1",fitfunc = svyglm,family=quasipoisson,data=dat[,-c(7:10,35,36)],design=nhanesdesign,m = 2,
                  numrs = 30,cores = 40,interactions = TRUE,criterion = int.p.val,minmax = 'min',
                 var4int = "diab",fixvar = c(adjust4b[c(1:2,4:5)],"ph")
                 )

tab1<-data.frame(print(fsafit_mh_2))

###########
#3-Way Int#
###########
fsafit_mh_3<-FSA(formula = "mh~1",fitfunc = svyglm,family=quasipoisson,data=dat[,-c(7:10,35,36)],design=nhanesdesign,m = 3,
                  numrs = 30,cores = 40,interactions = TRUE,criterion = int.p.val,minmax = 'min',
                 var4int = "diab",fixvar = c(adjust4b[c(1:2,4:5)],"ph")
                 )

tab2<-data.frame(print(fsafit_mh_3))

```

```{r, echo=FALSE}
kable(tab1, caption = "Mental Health Results for Two-Way Interactions")
dat$via<-as.factor(vita)
nhanesdesign<-svydesign(id=~SDMVPSU,
                        strata=~SDMVSTRA,
                        weights=~WTINT2YR,
                        nest=TRUE,
                        data=dat)

plot_model(svyglm(formula = "mh~RIDAGEYR+gender+edu+ses+ph+LBXVIA*diab",
                  design=nhanesdesign,family=quasipoisson,data=dat), type = "int",
           title = "LBXVIA*diab")
summary(svyglm(formula = "mh~RIDAGEYR+gender+edu+ses+ph+LBXVIA*diab",
                  design=nhanesdesign,family=quasipoisson,data=dat))
f1<-svyglm(formula = "mh~RIDAGEYR+gender+edu+ses+ph+LBXVIA*diab",
                  design=nhanesdesign,family=quasipoisson,data=dat)


```

<!-- \clearpage -->
<!-- \pagebreak -->

<!-- ```{r, echo=FALSE} -->
<!-- kable(tab2, caption = "Mental Health Results for Three-Way Interactions") -->
<!-- plot_model(svyglm(formula = "mh~RIDAGEYR+gender+edu+ses+BMXBMI+LBXVIE*LBXVIA*diab", -->
<!--                   design=nhanesdesign,family=quasipoisson,data=dat), type = "int", -->
<!--            title = "LBXVIE*LBDVIA*diab") -->
<!-- ``` -->

\clearpage
\pagebreak

## For Viatamin A
> "In contrast, elevated levels of retinyl stearate and retinyl palmitate were each estimated to enhance the risk for periodontitis particularly in smokers. The retinoids comprise a class of compounds related to Vitamin A. These compounds have been used to regulate epithelial cell growth, as well as playing a role in vision, regulation of cell proliferation and differentiation, growth of bone tissue, immune functions, and even activation of tumor suppressor genes70." 
From this paper [here](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6547714/)


\clearpage
\pagebreak

# Physical Health
```{r, echo=FALSE, include=FALSE}
##########################
#### Physical Health #######
##########################
nhanesdesign<-svydesign(id=~SDMVPSU,
                        strata=~SDMVSTRA,
                        weights=~WTINT2YR,
                        nest=TRUE,
                        data=dat)

null_model<-svyglm(formula = "ph~RIDAGEYR+gender+race+edu+ses+BMXBMI+mh",design=nhanesdesign,family=quasipoisson,data=dat)
#race  not significant will remove from null model and won't use in fixvar
null_model<-svyglm(formula = "ph~RIDAGEYR+gender+edu+ses+BMXBMI+mh+diab",design=nhanesdesign,family=quasipoisson,data=dat)



###########
#2-Way Int#
###########
fsafit_ph_2<-FSA(formula = "ph~1",fitfunc = svyglm,family=quasipoisson,data=dat[,-c(7:10,35,36)],design=nhanesdesign,m = 2,
                  numrs = 30,cores = 40,interactions = TRUE,criterion = int.p.val,minmax = 'min',
                 var4int = "diab",fixvar =  c(adjust4b[c(1:2,4:6)],"mh")
                 )

tab1<-data.frame(print(fsafit_ph_2))

###########
#3-Way Int#
###########
fsafit_ph_3<-FSA(formula = "ph~1",fitfunc = svyglm,family=quasipoisson,data=dat[,-c(7:10,35,36)],design=nhanesdesign,m = 3,
                  numrs = 30,cores = 40,interactions = TRUE,criterion = int.p.val,minmax = 'min',
                 var4int = "diab",fixvar =  c(adjust4b[c(1:2,4:6)],"mh")
                 )

tab2<-data.frame(print(fsafit_ph_3))

```

```{r, echo=FALSE}
kable(tab1, caption = "Physical Health Results for Two-Way Interactions")
plot_model(svyglm(formula = "ph~RIDAGEYR+gender+edu+ses+BMXBMI+mh+LBXVIE*diab",
                  design=nhanesdesign,family=quasipoisson,data=dat), type = "int",
           title = "LBXVIE*diab")
f2<-svyglm(formula = "ph~RIDAGEYR+gender+edu+ses+BMXBMI+mh+LBXVIE*diab",
                  design=nhanesdesign,family=quasipoisson,data=dat)
```

<!-- \clearpage -->
<!-- \pagebreak -->

<!-- ```{r, echo=FALSE} -->
<!-- kable(tab2, caption = "Physical Health Results for Three-Way Interactions") -->
<!-- plot_model(svyglm(formula = "ph~RIDAGEYR+gender+edu+ses+BMXBMI+LBXGTC*LBXALC*diab", -->
<!--                   design=nhanesdesign,family=quasipoisson,data=dat), type = "int", -->
<!--            title = "g-Tocopherol*AlphaCarotene*diab") -->
<!-- ``` -->

\clearpage
\pagebreak

## For AlphaCarotene, Vit E and Diabetic Complications
> "Our results further support the hypothesis that eating a diet rich in carotenoids, especially provitamin A carotenoids, might help prevent the development of type 2 diabetes in Japanese patients." [here](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4679813/)

> "Supplementation of vitamin E may be a valuable strategy for controlling diabetes complications and enhancing antioxidant capacity. The effects of other micronutrients should be further investigated in larger and well-designed trials to properly place these complementary therapies in clinical practice." From [here](https://dmsjournal.biomedcentral.com/articles/10.1186/s13098-018-0318-5)

